---
title: "IMDb Data Analysis"
output: pdf_document
---

# This document is designed to use data analysis to understand the movie ratings and how different service providers differ in the quality of movies on their platforms

## Loading the Packages

```{r, load-pkg}
library(tidyverse)
```

### Loading the file
```{r, load_file}
df_imdb <- read.csv("~/IMDb/Movies.csv", encoding = "UTF-8")
```

### Checking out the loaded file
```{r}
head(df_imdb,5)
colnames(df_imdb)
```

### Refining the dataset
```{r}
df_imdb <- df_imdb %>% 
  rename(c("Prime" = "Prime.Video", "Disney" = "Disney."))
df_imdb$Rotten.Tomatoes <- str_replace_all(df_imdb$Rotten.Tomatoes, "%", "")
df_imdb$Rotten.Tomatoes <- as.integer(df_imdb$Rotten.Tomatoes)
sum(is.na(df_imdb$Rotten.Tomatoes))
```

### Rechecking dataframe
```{r}
head(df_imdb, 5)
```

### Creating new data frame for data exploration
```{r}
df_imdb1 <- df_imdb %>% 
  select(Title, Year, IMDb, Rotten.Tomatoes, Netflix, Hulu, Prime, Disney) %>% 
  gather(Netflix:Disney, key = "provider", value = "has_movie")
```

### Checking new dataframe
```{r}
head(df_imdb1,5)
sprintf("Regular dataset has: %d", nrow(df_imdb))
sprintf("Gathered dataset has: %d", nrow(df_imdb1))
```

### Understading the data better
```{r}
df_imdb1 <- df_imdb1 %>% 
  filter(has_movie == 1)
nrow(df_imdb1)
write.csv(df_imdb1, "~/IMDb/Movies2.csv")
```

### Boxplot graph 

```{r, fig.height = 8, fig.width = 12}

df_imdb1 %>% 
  ggplot() + geom_boxplot(aes(x = provider, y = IMDb), na.rm = TRUE) + 
  labs(x = "provider", y = "IMDb ratings", title = "IMDb ratings across providers")

```

```{r, fig.height=8, fig.width = 12}
df_imdb1 %>% 
  ggplot() + 
  geom_boxplot(aes(x = provider, y = Rotten.Tomatoes), na.rm = TRUE) + 
  labs(x = "provider", y = "Rotten Tomatoes ratings") +
  ggtitle("Rotten Tomatoes ratings across providers")
```

### Graph of IMDb vs Rotten Tomatoes

```{r, fig.height = 8, fig.width=12}
model <- lm(IMDb ~ Rotten.Tomatoes, data = df_imdb)
coeff <- coef(model)
ggplot(df_imdb) +
  geom_point(aes(x = Rotten.Tomatoes, y = IMDb), color = "blue", na.rm = TRUE) +
  geom_abline(intercept = coeff[1], slope = coeff[2], color = "red") +
  theme_bw() + 
  labs(y = "IMDb", x = "Rotten Tomatoes", title = "IMDb ratings vs Rotten Tomatoes Ratings")
```

```{r}
df_imdb2 <- df_imdb %>% 
  mutate(new_rotten = as.double(Rotten.Tomatoes/10)) %>% 
  group_by(Year) %>% 
  summarize(mean_IMDb = mean(IMDb, na.rm = TRUE), mean_rotten = mean(new_rotten, na.rm=TRUE))
```

```{r, fig.height = 8, fig.width=12}
df_imdb2 %>% 
  gather(mean_IMDb:mean_rotten, key = "raters", value = "ratings") %>% 
  ggplot() + 
  geom_line(aes(x = Year, y = ratings, color = raters)) +
  labs(x = "year", y = "ratings", title = "How the movie ratings have differed over the years") +
  theme_light()
```

```{r, fig.height = 8, fig.width = 12}
ggplot(df_imdb2) +
  geom_line(aes(x = Year, y = mean_IMDb, color = "IMDb")) +
  geom_line(aes(x = Year, y = mean_rotten, color = "Rotten Tomatoes")) +
  labs(x = "year", y = "ratings", title = "How the movie ratings have differed over the years") +
  scale_colour_manual("", breaks = c("IMDb", "Rotten Tomatoes"), values = c("IMDb"="green", 
                                                                            "Rotten Tomatoes"="red"))+
  theme_bw()
```

## The Director with the best ratings
```{r}
df_imdb3 <- df_imdb %>% 
  select(Title, Directors, IMDb, Rotten.Tomatoes) %>% 
  separate(Directors, into = c("Dir1", "Dir2", "Dir3"), sep = ",") %>% 
  gather(Dir1:Dir3, key = "dir", value = "directors")
  
```

### Checking out new dataset
```{r, warning=FALSE}
df_imdb4 <- df_imdb3 %>% 
  filter(directors != "") %>% 
  group_by(directors) %>% 
  summarize(imdb = mean(IMDb, na.rm = TRUE), rotten = mean(Rotten.Tomatoes, na.rm = TRUE), count = n())
nrow(df_imdb4)
```


```{r}
df_imdb4 <- df_imdb4 %>% 
  mutate(rank = rank(desc(count), ties.method = "first")) %>% 
  arrange(rank)
```

```{r, fig.width = 12, fig.height = 8}
df_imdb4 %>% 
  filter(count >= 4) %>% 
  mutate(new_rank = rank(desc(imdb))) %>% 
  arrange(new_rank) %>% 
  filter(new_rank <= 10) %>%
  ggplot() +
  geom_bar(aes(x = reorder(directors, -imdb), y = imdb), fill = "blue", stat = "identity") +
  coord_cartesian(ylim = c(7, 10)) +
  labs(x = "Directors", y = "IMDb ratings", title = "Top 10 IMDB ratings with directors who have directed more than
       four movies") +
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

```{r, warning=FALSE}
df_imdb5 <- df_imdb %>% 
  separate(Genres, into = c("gen1", "gen2", "gen3", "gen4", "gen5","gen6", "gen7"), sep = ",") %>% 
  gather(gen1:gen7, key = "gen_num", value = "genre") %>% 
  filter(genre != "NA") %>% 
  select(Title,genre, IMDb, Rotten.Tomatoes)
nrow(df_imdb5)
```
```{r}
df_imdb6 <- df_imdb5 %>% 
  group_by(genre) %>% 
  summarize(imdb = mean(IMDb, na.rm = TRUE),rotten = mean(Rotten.Tomatoes, na.rm = TRUE), count = n()) %>% 
  mutate(rank = rank(desc(count), ties.method = "first")) %>% 
  arrange(rank)
```


```{r, fig.width = 12, fig.height = 8}
df_imdb6 %>% 
  mutate(new_rank = rank(desc(imdb))) %>% 
  filter(new_rank <= 10) %>% 
  ggplot() +
  geom_bar(aes(x = reorder(genre, -imdb), y = imdb), fill = "purple", stat = "identity") +
  coord_cartesian(ylim = c(5, 9)) +
  labs(x = "Genres", y = "IMDb ratings", title = "Top 10 Genres with the highest IMDb ratings") +
  theme(axis.text.x=element_text(angle=45, hjust=1))
  
```



